
---
title: "STAT 1910 Project Data Proposal"
author: "team 11: [individual student names redacted]"
date: "Last updated on [date redacted]"
output:
  html_document:
    fig_caption: yes
    theme: lumen
    toc: yes
    toc_depth: 2
    df_print: kable
    toc_float:
      collapsed: no
---

```{r, message=FALSE}
# Load all packages here:
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)

# Set seed value of random number generator to get "replicable" random numbers.
# The choice of seed value of 76 was an arbitrary one on my part.
set.seed(9)
```


# Big-picture

## Research question

_What is your research question?_


 As students, we were interested in a topic that pertains to education, and we are all from different backgrounds, which brought up many issues on how race and graduation are linked. Thus, we have decided to determine and comprehend if a student's race affects their graduation and what factors into it. 


## Description of data

_Please give a very short description of the data set along with it's original source._

Finding the data was an essential component, as we wanted a city where there are different ethnicities or a town that is known to be very diverse. We chose New York and got the data on high school graduation rates during 2017 from the Department of Education. Also, we focused on the cohort year 2017, which was pretty recent, and the data contains the percentage of dropouts for each Race in that same year. 

The data was downloaded from [New York Department of Education site](https://infohub.nyced.org/reports/academics/graduation-results)

## Load data into R

_Include the code to load your data here. If your data is not confidential nor is it private in nature, consider publishing it as a `.csv` file on Google Sheets as in the code chunk below; instructions on how to do this are in Steps 1-6 [here](https://twitter.com/rudeboybert/status/1055821846539591680). If the data shouldn't be published online, then please submit the spreadsheet file on Moodle._

```{r, message=FALSE}
grade <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSBf2AKc_0S8dDMzr7CfK7LPRS3g0GRtq-B7wWIia-j8kw0KPECR2lKHR05lsP98rqNnWSNbdnj4FRq/pub?output=csv")
```


## Clean variable names

_Piping your data frame into the `clean_names()` function from the `janitor` package will clean your variable names, making them easier to work with._

```{r}
grade <- grade %>% 
  clean_names()

```


## Explore your data

_Be sure to explore your data. Note that `eval=FALSE` is set so that R Markdown doesn't "evaluate" this code chunk, i.e. it will ignore it in the ultimate `.html` report. You should run this code on your own, but not in the ultimate `.html` report._

```{r, eval = FALSE}
glimpse(grade)
```



***



# Variables

## Identification variable

_What is your identification (ID) variable (if you have one)?_

Our identification variable will be the school's name since we will compare the different schools and their graduation rate for each Race. Individuals will be able to identify each one.

## Outcome variable

_What is your outcome variable $y$? What are its units of measurement?_

The outcome variable is the percentage of high school graduates for each Race, and its units of measurement are the percentage. 


## Numerical explantory variable

_What is your numeric explanatory variable? What are its units of measurement?_

The numerical explanatory variable is the percent dropouts, and its unit of measurement is percentage points between 0-100. This variable would indicate the number of students who dropped out from each Race.

## Categorical explantory variable

_What is your categorical explanatory variable? Does it have between 3 and 5 levels. Please list the different levels._

Our categorical explanatory variable would be the student's Race and has six levels: White, Black, Asian, Hispanic, Multi-Racial and Native American. 

***


# Rows/observations

## Observational units

_What is the observational unit of your data? In other words, what does each row in your data represent?_

Each row in our data set will be a high school in New York City. 


## Sample size

_How many rows/cases are in the data i.e. what is the sample size? Is the sample size at least 50?_

At the moment, there are 131,610 rows and nine columns. Although, we will decrease our number of rows to schools with all six races, not just 2 or 3.

***


# Preview of data

## Pare down variables

_`select()` the following variables **in this order** and drop all others. Eliminating all unnecessary variables will making visually exploring the raw values less taxing mentally, as we'll have less data to look at._

1. _The identification variable_
1. _The outcome variable $y$_
1. _The numerical explanatory variable_
1. _The categorical explanatory variable_
1. _Optional: any other variable you find interesting_


```{r}
grade <- grade %>% 
  select(school_name, percent_grads   , percent_dropout  , category)
glimpse(grade)

```


## Preview data

_Display a random sample of 5 rows of your data frame by piping it into the `sample_n(5)` function from the `dplyr` package . You'll get the same 5 rows everytime you knit this document and hence replicable results because we set the seed value of the random number generator in the first code chunk above._

```{r}
grade %>% 
  sample_n(5)
```




# Author Statement

## Individual Roles

_Describe the role (Facilitator, Recorder, Reporter, Time Keeper, Checker) and level of contribution of each group member during this project phase_ 

All group members contributed equally to this submission. 

* \[Redacted] acted as the  as the facilitator. 
* \[Redacted] acted as the Recorder. 
* \[Redacted] acted as the Reporter. 
* \[Redacted] acted as the Time Keeper.
* \[Redacted] acted as the Checker. 


## Individual Contribution

_If your group were to earn 5 points on this submission, how should those points be shared across your group members?_


* \[Redacted]:  1 points.
* \[Redacted]: 1 points.
* \[Redacted]:  1 points.
* \[Redacted]: 1 points.
* \[Redacted]: 1 points.


