---
title: "STAT 1910 Project Data Proposal - Diamonds Dataset"
author: "Team 04: [individual student names redacted]"
date: "Last updated on [date redacted]"
output:
  html_document:
    fig_caption: yes
    theme: lumen
    toc: yes
    toc_depth: 2
    df_print: kable
    toc_float:
      collapsed: no
---

# Setup

## Load packages

```{r, message=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)
```

## Set seed value

Our choice of seed value is 42 - "The answer to life, the universe, and everything" - [*Douglas Adams' The Hitchhiker's Guide to the Galaxy*](https://www.techtarget.com/whatis/definition/42-h2g2-meaning-of-life-The-Hitchhikers-Guide-to-the-Galaxy)

```{r message=FALSE}
set.seed(42)
```

# Big-picture

## Research question

In this project, we are trying to understand what attributes of diamonds influence their price the most.

## Description of data

This data set contains the prices and other attributes of 53,940 diamonds. The data was originally downloaded from Kaggle [here](https://www.kaggle.com/datasets/shivam2503/diamonds) and was modified to include `id` column.

## Load data into R

```{r, message=FALSE}
diamonds <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vQdSy4ef3CfFtHvj7BjzMHNXhLdpFwaRBgBGc6was-iiM4F0ILQQolGeUBnip9K71novyrNuV3bGaVh/pub?gid=1644666616&single=true&output=csv")
```

## Clean variable names

```{r}
diamonds <- clean_names(diamonds)
```

## Explore your data

```{r, eval = FALSE}
glimpse(diamonds)
```

------------------------------------------------------------------------

# Variables

## Identification variable

Our identification variable is `id`. Which is just the id number of each diamond recorded.

## Outcome variable

Our outcome (response/dependent) variable is the `price` of the diamond in US dollars (\$326 -- \$18,823).

## Numerical explanatory variable

Our numerical explanatory variable is `carat` which represents the weight of the diamond in carats (0.2 - 5.01).

## Categorical explanatory variable

Our categorical explanatory variable is `cut` which is the quality of the cut. It has 5 levels (Fair, Good, Very Good, Premium, Ideal).

------------------------------------------------------------------------

# Rows/observations

## Observational units

Each row represents a unique diamond.

## Sample size

There are 53,940 rows (diamonds) in the data set.

------------------------------------------------------------------------

# Preview of data

## Pare down variables

1.  *The identification variable* $id$
2.  *The outcome variable* $price$
3.  *The numerical explanatory variable* $carat$
4.  *The categorical explanatory variable* $cut$

```{r}
diamonds <- diamonds %>%
  select(id, price, carat, cut)
```

## Preview data

```{r}
sample_n(diamonds, 5)
```
