
---
title: "STAT 1910 Project Data Proposal"
author: "Team 07: [individual student names redacted]"
date: "Last updated on [date redacted]"
output:
  html_document:
    fig_caption: yes
    theme: lumen
    toc: yes
    toc_depth: 2
    df_print: kable
    toc_float:
      collapsed: no
---

```{r, message=FALSE}
# Load all packages here:
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)

# Set seed value of random number generator to get "replicable" random numbers.
# The choice of seed value of 76 was an arbitrary one on my part.
set.seed(76)
```


# Big-picture

## Research question

_What is your research question?_

This study aims to understand how and whether rotten tomatoes ratings will influence the profitability of Hollywood movies. To be specific the question is whether the ratings and genre of movies will affect the percentage of profitability for Hollywood movies. 

## Description of data

_Please give a very short description of the data set along with it's original source._

We have data from different genres of various profitable Hollywood movies in 2011 and its corresponding variables that describe it. The data set contains the rotten tomatoes ratings and profitability of the movies, both of which are measured in percentage. The data was obtained from the orignial source itself which was the "information is beautiful" website, in which you can access it through a title named "Hollywood Budgets". This is the link to access it: https://informationisbeautiful.net/2012/hollywood-budgets-a-5000-data-viz-challenge/

## Load data into R

_Include the code to load your data here. If your data is not confidential nor is it private in nature, consider publishing it as a `.csv` file on Google Sheets as in the code chunk below; instructions on how to do this are in Steps 1-6 [here](https://twitter.com/rudeboybert/status/1055821846539591680). If the data shouldn't be published online, then please submit the spreadsheet file on Moodle._

```{r, message=FALSE}
movies_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vT-DqFII3DOoUeCKuVe-gZASgKWJPgM7JnHhhUUGgcXpdLjx4KmYnUZiR1kQ0mjOg/pub?gid=1286148029&single=true&output=csv")
```


## Clean variable names

_Piping your data frame into the `clean_names()` function from the `janitor` package will clean your variable names, making them easier to work with._

```{r}
movies_data <- movies_data %>% clean_names()
```


## Explore your data

_Be sure to explore your data. Note that `eval=FALSE` is set so that R Markdown doesn't "evaluate" this code chunk, i.e. it will ignore it in the ultimate `.html` report. You should run this code on your own, but not in the ultimate `.html` report._

```{r, eval = FALSE}
glimpse(movies_data)
```



***



# Variables

## Identification variable

_What is your identification (ID) variable (if you have one)?_

The identification variable is "film". It helps to identify the names of the movies.

## Outcome variable

_What is your outcome variable $y$? What are its units of measurement?_

The outcome variable is called "profitability". The profitability of the movies are measured in percentage within the range of 0 - 7000.

## Numerical explantory variable

_What is your numeric explanatory variable? What are its units of measurement?_

The numerical explanatory variable is called "rotten_tomatoes". It is measured in percentage within the range of 0 - 100. It represents the ratings of the movies.

## Categorical explantory variable

_What is your categorical explanatory variable? Does it have between 3 and 5 levels. Please list the different levels._

The categorical explanatory variable is "genre". The categorical values in genre are a total of 9 types. We had to minimize the levels of genres to 5. The values like "Fantasy", "Adventure", "Romance", "Animation", and "Thriller" will be grouped into one value named "Other" which will be solved in the later stages of data wrangling. Therefore, the 5 categorical values will become "Action", "Comedy", "Drama", "Horror" and "Other".

***


# Rows/observations

## Observational units

_What is the observational unit of your data? In other words, what does each row in your data represent?_

Each row in the data set is a Hollywood movie with its corresponding details like its genre, audience score, profitability, rotten tomatoes, budget, etc.

## Sample size

_How many rows/cases are in the data i.e. what is the sample size? Is the sample size at least 50?_

In our original data set there are 143 rows but after eliminating missing values there will be 139 rows, which will be done through data wrangling in the later stages.

***


# Preview of data

## Pare down variables

_`select()` the following variables **in this order** and drop all others. Eliminating all unnecessary variables will making visually exploring the raw values less taxing mentally, as we'll have less data to look at._

1. _The identification variable_
1. _The outcome variable $y$_
1. _The numerical explanatory variable_
1. _The categorical explanatory variable_
1. _Optional: any other variable you find interesting_


```{r}
movies_data <- movies_data %>% select(film, profitability, rotten_tomatoes, genre, lead_studio)
```


## Preview data

_Display a random sample of 5 rows of your data frame by piping it into the `sample_n(5)` function from the `dplyr` package . You'll get the same 5 rows everytime you knit this document and hence replicable results because we set the seed value of the random number generator in the first code chunk above._

```{r}
movies_data %>% sample_n(5)
```




# Author Statement

## Individual Roles

_Describe the role (Facilitator, Recorder, Reporter, Time Keeper, Checker) and level of contribution of each group member during this project phase_ 

\[Redacted] did not contribute to this submission. \[Redacted] participated in the first discussions, but made only a small contribution to this project. The other 3 group members contributed equally. 

* \[Redacted] acted as the  as the facilitator. 
* \[Redacted] acted as the Recorder. 
* \[Redacted] acted as the Reporter. 
* \[Redacted] was assigned the role of the Time Keeper.
* \[Redacted] was assigned the role of the Checker. 


## Individual Contribution

_If your group were to earn 5 points on this submission, how should those points be shared across your group members?_


* \[Redacted]:  0 points.
* \[Redacted]:  0.5 points.
* \[Redacted]: 1.5 points.
* \[Redacted]: 1.5 points.
* \[Redacted]: 1.5 points.


