
---
title: "STAT 1910 Project Data Proposal"
author: "Group 06: [individual student names redacted]"
date: "Last updated on [date redacted]"
output:
  html_document:
    fig_caption: yes
    theme: lumen
    toc: yes
    toc_depth: 2
    df_print: kable
    toc_float:
      collapsed: no
---

```{r, message=FALSE}
# Load all packages here:
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)

# Set seed value of random number generator to get "replicable" random numbers.
# The choice of seed value of 76 was an arbitrary one on my part.
set.seed(76)
```


# Big-picture

## Research question

_What is your research question?_

How age and weekday alcohol consumption affects grades.


## Description of data

_Please give a very short description of the data set along with it's original source._

The data set studies two Portuguese schools, recording various pieces of information. We are focusing on the age, weekday alcohol consumption, and how they correlate to their grades throughout the year. 

https://www.kaggle.com/datasets/whenamancodes/alcohol-effects-on-study


## Load data into R

_Include the code to load your data here. If your data is not confidential nor is it private in nature, consider publishing it as a `.csv` file on Google Sheets as in the code chunk below; instructions on how to do this are in Steps 1-6 [here](https://twitter.com/rudeboybert/status/1055821846539591680). If the data shouldn't be published online, then please submit the spreadsheet file on Moodle._

```{r, message=FALSE}
dataset <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSL7hmQ0GKusiQrdLUSLT227_g3wU4DXxv1Hjff0vUWwCqLDWGYv7OYtIGQuRSHwiSMoTqE7HLQvJi8/pub?output=csv")
```


## Clean variable names

_Piping your data frame into the `clean_names()` function from the `janitor` package will clean your variable names, making them easier to work with._

```{r}
dataset <- dataset %>%
  clean_names()
```


## Explore your data

_Be sure to explore your data. Note that `eval=FALSE` is set so that R Markdown doesn't "evaluate" this code chunk, i.e. it will ignore it in the ultimate `.html` report. You should run this code on your own, but not in the ultimate `.html` report._

```{r, eval = TRUE}
glimpse(dataset)
```



***



# Variables

## Identification variable

_What is your identification (ID) variable (if you have one)?_

NA

## Outcome variable

_What is your outcome variable $y$? What are its units of measurement?_

The outcome variable is G3, which represents the final grade. 

## Numerical explantory variable

_What is your numeric explanatory variable? What are its units of measurement?_

The numerical variable is age, in years. 

## Categorical explantory variable

_What is your categorical explanatory variable? Does it have between 3 and 5 levels. Please list the different levels._

The categorical variable is Dalc, workday alcohol consumption. Levels 1 - very low, through 5 - very high. 

***


# Rows/observations

## Observational units

_What is the observational unit of your data? In other words, what does each row in your data represent?_

Each row represents an individual from the Portuguese schools. 

## Sample size

_How many rows/cases are in the data i.e. what is the sample size? Is the sample size at least 50?_

There are 395 rows/cases.

***


# Preview of data

## Pare down variables

_`select()` the following variables **in this order** and drop all others. Eliminating all unnecessary variables will making visually exploring the raw values less taxing mentally, as we'll have less data to look at._

1. _The identification variable_
1. _The outcome variable $y$_
1. _The numerical explanatory variable_
1. _The categorical explanatory variable_
1. _Optional: any other variable you find interesting_


```{r}
dataset <- dataset %>%
  select(g3, age, dalc, health)
```


## Preview data

_Display a random sample of 5 rows of your data frame by piping it into the `sample_n(5)` function from the `dplyr` package . You'll get the same 5 rows everytime you knit this document and hence replicable results because we set the seed value of the random number generator in the first code chunk above._

```{r}
dataset %>% 
  sample_n(5)
```





